---
title: "Monte Carlo Simulation for Model Evaluation"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

### Simulation with known abundance

```{r}
#| label: sim

## simulating data
require(tidyverse)
ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
mean(ytrue == 0)  
y <- ytrue[ytrue>0]  
n <- length(y)  
table(ytrue) %>% plot
```

### `stan` model:

```{r}
#| label: stan_model

mod <- "  
data {  
int<lower = 1> n; // total number of species  
array[n] int<lower = 0> y; // catch per species  
}

parameters {  
real log_mu;  
real<lower = 0> r;  
}

transformed parameters {  
real<lower=0,upper=1> p0;  
real<lower=0> mu;  
mu = exp(log_mu);  
p0 = (r/(mu+r))^r;  
}

model {  
log_mu ~ normal(0,5);  
r ~ cauchy(0,2);

target += neg_binomial_2_log_lpmf(y | log_mu, r) - log1m(p0);  
}

generated quantities {  
real N; // estimated total number of species  
N = n/(1-p0);  
}  
"
  
fit_mod <- stan_model(model_code = mod)  

fit <- sampling(fit_mod, data=list(n=n, y=y), chains=8)
print(fit)

```

### Sampling distribution

Simulation to estimate the sampling distribution of N

```{r}
#| label: sampling_distribution
 
nsim <- 100
fitN <- rv()
fit_pnpmle <- fit_jkknf <- fit_ace <- fit_lowerbound <- 
  fit_covdur <- fit_unpmle <- fit_pcg <- list()
for (i in 1:nsim){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (nsim+1):(2*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (2*nsim+1):(3*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (3*nsim+1):(4*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (4*nsim+1):(5*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (5*nsim+1):(6*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (6*nsim+1):(7*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (7*nsim+1):(8*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (8*nsim+1):(9*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}

for (i in (9*nsim+1):(10*nsim)){
  ytrue <- rnbinom(n = 60, size = 10, prob = 0.85)  
  mean(ytrue == 0)  
  y <- ytrue[ytrue>0]  
  n <- length(y)
  n_j <- table(y)
  y.frame <- matrix(c(1:length(n_j), n_j), ncol=2, byrow=F)
  fit <- sampling(fit_mod, data=list(n=n, y=y), 
                  iter=niters, thin=nthin, chains=nchains)
  fitN[i] <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))[5]
  fit_pnpmle[[i]] <- pnpmle(y.frame, t=15, C=1, b=200)
  fit_jkknf[[i]] <- jackknife(y.frame, k=5)
  fit_ace[[i]] <- ChaoLee1992(y.frame, t=5)
  fit_lowerbound[[i]] <- chao1984(y.frame)
  fit_covdur[[i]] <- ChaoBunge(y.frame, t=5)
  fit_unpmle[[i]] <- unpmle(y.frame, t=5, C=1, b=200)
  ##fit_pcg[[i]] <- pcg(y.frame, t=10, C=1, b=200)
  print(i)
}
```

### Summarizing simulation results

Summarizing simulation results 1. Graphical comparison of the estimated CIs (95% credible intervals or confidence intervals) 2. Summary statistics of the estimation errors (estimated - true): - absolute deviance (accuracy) - mean residuals (bias) - mean squared errors (square differences)

```{r}
#| label: summarizing

oo <- sample(1000, 100, F)
tikz(file="div_sims.tex", height=6, width=6.25, standAlone = F)
par(mfrow=c(1,2), mar=c(4, 1, 2, 1), mgp=c(1.5,0.25,0), tck=-0.01)
mlplot(fitN[oo], xlim=c(40, 150), cex=0.5, axes=F, 
       xlab="Estimated Richness", ylab="simulations",
       sub="Wang and Lindsay (2008)")
abline(v=60, col="red")
axis(1, at=seq(30, 160, 20))
axis(3, at=c(60), tck=0.02, col="red")
axis(3, at=c(35,250), tck=0.02)
text(x=60, y=-6.5, "True richness")
for (i in 1:100){
  temp <- fit_pnpmle[[oo[i]]]
  points(x=temp$Nhat, y=i+0.5, col=rgb(0.5,0.5,0.5,0.5), pch=16, cex=0.5)
  segments(x0=temp$CI[1], x1=temp$CI[2], y0=i+0.5, y1=i+0.5, col=rgb(0.5,0.5,0.5,0.5))
}

mlplot(fitN[oo], xlim=c(40, 150), cex=0.5, axes=F, 
       xlab="Estimated Richness", ylab="simulations", 
       sub="Chao and Lee (1992)")
abline(v=60, col="red")
axis(1, at=seq(30, 160, 20))
axis(3, at=c(60), tck=0.02, col="red")
axis(3, at=c(35,250), tck=0.02)
text(x=60, y=-6.5, "True richness")
for (i in 1:100){
  temp <- fit_ace[[oo[i]]]
  points(x=temp$Nhat[2], y=i+0.5, col=rgb(0.5,0.5,0.5,0.5), pch=16, cex=0.5)
  segments(x0=temp$CI[2,1], x1=temp$CI[2,2], y0=i+0.5, y1=i+0.5, col=rgb(0.5,0.5,0.5,0.5))
}

dev.off()

save(fitN, fit_pnpmle, fit_jkknf, fit_ace, fit_lowerbound,
     fit_covdur, fit_unpmle, file="sim_raw.RData")

## calculating mead differences
N_pnpmle <- N_jkknf <- N_ace <- N_lowerbound <- N_unpmle <- N_covdur <- CIw_pnpmle <- CIw_jkknf <- CIw_ace <- CIw_lowerbound <- CIw_unpmle <- CIw_covdur <-numeric() 

for (i in 1:(nsim*10)){
  N_pnpmle[i] <- fit_pnpmle[[i]]$Nhat
  N_jkknf[i] <- fit_jkknf[[i]]$Nhat
  N_ace[i] <- fit_ace[[i]]$Nhat[1]
  N_lowerbound[i] <- fit_lowerbound[[i]]$Nhat
  N_unpmle[i] <- fit_unpmle[[i]]$Nhat
  N_covdur[i] <- fit_covdur[[i]]$Nhat
  CIw_pnpmle[i] <- diff(fit_pnpmle[[i]]$CI[1,])
  CIw_jkknf[i] <- diff(fit_jkknf[[i]]$CI[1,])
  CIw_ace[i] <- diff(fit_ace[[i]]$CI[2,])
  CIw_lowerbound[i] <- diff(fit_lowerbound[[i]]$CI[1,])
  CIw_unpmle[i] <- diff(fit_unpmle[[i]]$CI[1,])
  CIw_covdur[i] <- diff(fit_covdur[[i]]$CI[1,])
}

CIw_Bayes <- apply(summary(fitN)[, c(4,8)], 1, diff)

CIw_all <- c(rvsims(CIw_Bayes),  rvsims(CIw_pnpmle), rvsims(CIw_jkknf), rvsims(CIw_ace), rvsims(CIw_lowerbound), rvsims(CIw_unpmle), rvsims(CIw_covdur))

names(CIw_all) <- c("Bayes", "pnpmle", "jackknife", "ChaoLee1992","chao1984", "ChaoBungle", "unpmle")

tikz(file="CIws.tex", width=4, height=3., standAlone = F)
par(mar=c(3, 3, 1, 1), mgp=c(1.25, 0.125, 0), tck=0.01)
mlplot(CIw_all, xlim=c(4, 100), xlab="Confidence/credible Interval Widths")
dev.off()

rev_table <- cbind(c(mean(abs(summary(fitN)$mean - 60)),
mean(abs(N_pnpmle-60)),
mean(abs(N_jkknf-60)),
mean(abs(N_ace-60)),
mean(abs(N_lowerbound-60)),
mean(abs(N_unpmle-60)),
mean(abs(N_covdur-60))), 

c(mean(summary(fitN)$mean - 60),
mean(N_pnpmle-60),
mean(N_jkknf-60),
mean(N_ace-60),
mean(N_lowerbound-60),
mean(N_unpmle-60),
mean(N_covdur-60)),

sqrt(c(mean((summary(fitN)$mean - 60)^2),
mean((N_pnpmle-60)^2),
mean((N_jkknf-60)^2),
mean((N_ace-60)^2),
mean((N_lowerbound-60)^2),
mean((N_unpmle-60)^2),
mean((N_covdur-60)^2))))

colnames(rev_table) <- c("Accuracy", "Bias", "RMSE")

rownames(rev_table) <- c("Bayes", "pnpmle", "jackknife", "ChaoLee1992","chao1984", "ChaoBungle", "unpmle")
save(N_pnpmle, N_jkknf, N_ace, N_lowerbound, N_unpmle, N_covdur,
     file = "sim_results.RData")
write.table(round(rev_table, 3), file="rev_table.tex", sep="&")
```
